```{r child='boxStyles.Rmd'}
```
<p class=box>
$T_{0}^{\ddagger}$= 288.15 K, reference temperature  
$\lambda_{a}^{\ddagger}$ \sindex[lis]{lambdaa@$\lambda_{a}$= tropospheric lapse rate, standard atmosphere}=
-0.0065 $^{\circ}$C per geopotential meter = the lapse rate\index{International Standard Atmosphere!lapse rate} for the troposphere$^{\ddagger}$  
$p$ = measured static (ambient) pressure, hPa, usually from PSXC  
$p_{0}^{\ddagger}$\sindex[lis]{P0star@$p_{0}^{\ddagger}$= reference pressure for zero altitude,
ISA} = 1013.25 hPa, reference pressure for PALT=0 $^{\ddagger}$  
$M_{d}^{\ddagger}$ = 28.9644 kg/kmol = molecular weight of dry
air, ISA definition $^{\ddagger}$  
$g^{\ddagger}$ = 9.80665 m\ s$^{-2}$, acceleration of gravity $^{\ddagger}$  
$R_{0}^{\ddagger}$ = universal gas constant, defined$^{\ddagger}$
as 8.31432$\times10^{3}$ J\ kmol^-1^\ K^-1^      
$z_{T}^{\ddagger}$ = altitude of the ISA tropopause = 11,000 m $^{\ddagger}$  
$x=-R_{0}^{\ddagger}\lambda_{a}^{\ddagger}/(M_{d}^{\ddagger}g^{\ddagger})$
$\approx$ 0.1902632 (dimensionless)\footnote{This is the value, rounded to seven significant figures, that is used for data processing.}  
</p>
<p class=box2>

For pressure > 226.3206 hPa (equivalent to a pressure altitude < $z_{T}$):
\begin{equation}
\mathrm{PALT}=-\left(\frac{T_{0}^{\ddagger}}{\lambda^{\ddagger}}\right)\left(1-\left(\frac{p}{p_{0}^{\ddagger}}\right)^{x}\right)
(\#eq:PALT1)
\end{equation}  
otherwise, if $T_{T}$\sindex[lis]{Tt@$T_{T}$= temperature at the ISA tropopause}
and $p_{T}$\sindex[lis]{pT@$p_{T}$= pressure at the ISA tropopause}
are respectively the temperature and pressure at the altitude\index{International Standard Atmosphere!tropopause} $z_{T}$:  
$$T_{T}=T_{0}+\lambda^{\ddagger}z_{T}^{\ddagger}=216.65\,\mathrm{K}$$
$$p_{T}=p_{0}^{\ddagger}\Bigl(\frac{T_{0}^{\ddagger}}{T_{T}}\Bigr)^{\frac{g^{\ddagger}M_{d}^{\ddagger}}{\lambda^{\ddagger}R_{0}^{\ddagger}}}=226.3206\,\mathrm{hPa}$$
\begin{equation}
\mathrm{PALT}=z_{T}^{\ddagger}+\frac{R_{0}^{\ddagger}T_{T}}{g^{\ddagger}M_{d}^{\ddagger}}\ln\left(\frac{p_{T}}{p}\right)
(\#eq:PALT2)
\end{equation}  
which, after conversion from natural to base-10 logarithm, is coded
to be equivalent to the following:  
</p>
```{r class.source="box-out", echo=TRUE, include=TRUE, eval=FALSE, highlight=FALSE}
// transition pressure at the assumed ISA tropopause:
#define ISAP1 226.3206
// reference pressure for standard atmosphere:
#define ISAP0 1013.25
if (psxc > ISAP1):
 palt = 44330.77 * (1.0 - pow(psxc / ISAP0, 0.1902632));
else:
 palt = 11000.0 + 14602.12 * log10(ISAP1 / psxc);
```
