---
title: 'RAF Technical Note: Processing Algorithms'
author: "RAF RSIG"
date: "May 2022"
documentclass: book
site: bookdown::bookdown_site
always_allow_html: yes
---

```{r init, include=FALSE}
options(tinytex.verbose = TRUE)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
library(magrittr)
PToDo <- TRUE # Set TRUE to include ToDo indicators.
```

# Preface {-#index}

Algorithms Used to Produce Data Products from Research Aircraft;  
also Definitions of Variables in Archived Data Files  

Previous PDF version of this document: [Old PDF Version](www/ProcessingAlgorithms.pdf).    
Download the current version in a browser for functional URL links. The HTML version generated by the .Rmd-format version can be viewed [here](https://williamcooper.github.io/ProcessingAlgorithms) and the RMarkdown code is archived [here](https://github.com/WilliamCooper/ProcessingAlgorithms/tree/Rmd/).  Many changes, mostly cosmetic but some substantive, have been
made to the PDF-format version of February 2019, which was generated using the program LyX
and the
files still available [here](https://github.com/WilliamCooper/ProcessingAlgorithms/tree/lyx/).

Displays involving mathematical expressions rely on MathJax for formatting, and some delay may be present between initially loading the HTML pages and completing this formatting.  

```{r child='Children/boxStyles.Rmd'}

```

```{r child='Children/Section1.Rmd'}
```

```{r child='Children/Section2.Rmd'}
```

```{r child='Children/Section3.Rmd'}
```

```{r child='Children/Section4.Rmd'}
```

```{r child='Children/Section5.Rmd'}
```

```{r child='Children/Section6.Rmd'}
```

```{r child='Children/Section7.Rmd'}
```

```{r child='Children/Section8.Rmd'}
```

```{r child='Children/Section9.Rmd'}
```

```{r child='Children/Section10.Rmd'}
```

```{r child='Children/Appendices.Rmd'}
```
 
 ```{r child='Children/Index.Rmd'}
 ```

