---
title: 'RAF Technical Note: Processing Algorithms'
author: "RAF RSIG"
date: "March 2021"
output:
  html_document:
    df_print: paged
    keep_tex: yes
    split_by: chapter+number
    toc: true
    toc_depth: 2
    toc_float: true
  bookdown::gitbook:
    lib_dir: assets
    split_by: chapter+number
    keep_tex: yes
    # includes:
    #   in_header: columns.tex
  html_document2:
    keep_tex: yes
    toc: true
    toc_depth: 2
    toc_float: true
    includes:
      in_header: columns.tex
    config:
      toolbar:
        position: static
documentclass: book
site: bookdown::bookdown_site
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
```

# Preface {-#index}

Algorithms Used to Produce Data Products from Research Aircraft;  
also Definitions of Variables in Archived Data Files  

Previous PDF version of this document: [Old PDF Version](https://github.com/NCAR/aircraft_ProcessingAlgorithms/blob/master/ProcessingAlgorithms.pdf) <br />  
Download the current version in a browser for functional URL links. The HTML version generated by the .Rmd-format version can be accessed [here](https://htmlpreview.github.io/?https://github.com/WilliamCooper/ProcessingAlgorithms/blob/Rmd/index.html) and the RMarkdown code is archived [here](https://github.com/WilliamCooper/ProcessingAlgorithms/tree/Rmd/).  Although the new
date on the RMarkdown version is March 2021, only a few (mostly cosmetic) changes have been
made to the PDF-format version of February 2019, which was generated using the program LyX
and the
files still available [here](https://github.com/WilliamCooper/ProcessingAlgorithms/tree/master/).

```{r child='Children/boxStyles.Rmd'}

```

```{r child='Children/Section1.Rmd'}
```

```{r child='Children/Section2.Rmd'}
```

```{r child='Children/Section3.Rmd'}
```

```{r child='Children/Section4.Rmd'}
```

```{r child='Children/Section5.Rmd'}
```

```{r child='Children/Section6.Rmd'}
```

```{r child='Children/Section7.Rmd'}
```

```{r child='Children/Section8.Rmd'}
```

```{r child='Children/Section9.Rmd'}
```

```{r child='Children/Section10.Rmd'}
```

```{r child='Children/Appendices.Rmd'}
```
 
 ```{r child='Children/Index.Rmd'}
 ```

