---
title: "Table of Variables"
author: "William A Cooper"
date: "`r Sys.Date()`"
output: html_document
---

```{r include = FALSE, echo = FALSE}
options(tinytex.verbose = TRUE)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
library(Ranadu)
library(magrittr)
library(kableExtra)
```

The following table lists variables from the netCDF archives, and tries to classify them into these types:  

* pa: in the ProcessingAlgorithms document  
* a:  should be added to the ProcessingAlgorithms document, at least in the appendix.  
* st: variables that provide status or housekeeping information. Some may be included in PA.  
* sp: special-use variables from, e.g., user or non-standard instruments.  
* v: an alternate version of a variable otherwise described.  
* npa: not in PA document, and should not be (e.g., JHWRP variables from UW)  
* undefined: unable to classify because insufficient info in the netCDF files, or just not classified yet.  




```{r tos, include=FALSE, echo=FALSE}
SymbolTable <- data.frame("abbreviation-or-symbol"=c('º', 'ft', 'mb', 'ppmv', 'ppbv', 'pptv', 'n mi', 'kt'),
                          "definition$^a$" = c('degree, angle measurement ≡ $\\pi/180$', 'foot ≡ 0.3048 m',
                                         'millibar ≡ 100 Pa ≡ 1 hPa',
                                         'parts per million by volume (see item 4 above)',
                                         'parts per billion ($10^9$) by volume (see item 4 above)',
                                         'parts per trillion by volume (see item 4 above)',
                                         'nautical mile ≡1852 m',
                                         'knot (n mi/hour) ≡ (1852/3600) m/s = 0.514444… m/s'))
names(SymbolTable) <- c('abbreviation or symbol', 'definition$^a$')
```

```{r tov, include = TRUE}
load('DVALL.Rdata')  # Load the list of variables in DV
# Display only 'a' type:
DVS <- DV[DV$type == 'a',]
knitr::kable(DVS[, -c(2:4, 6)], row.names=FALSE, align=c('cc')) %>% kable_styling(bootstrap_options = 'bordered') %>%
  add_footnote("a footnote", notation="alphabet") %>%
  scroll_box(width='90%')
```

