# Aerosol Particle Measurements {#aerosol-particle-measurements}

```{r child='boxStyles.Rmd'}
```
<!-- <h2 id="condensation-nucleus-counters">Condensation Nucleus Counters</h2> -->
## Condensation Nucleus Counter {#condensation-nucleus-counter}

RAF uses two modified TSI, Inc. [condensation nucleus counters](http://www.eol.ucar.edu/instruments/condensation-nucleus-counter-water-or-butanol) to measure the total concentration of fine or ultrafine particles in the atmosphere, a 3760A using n-butyl alcohol and a water-based 3786 WCN (water condensation nucleus) counter. Both are sensitive to particles in the approximate diameter range from $0.010–3\,\mu\mathrm{m}$.  

#### CN Counter Inlet Pressure (hPa): PCN, PACN {-#pcn}

*The absolute pressure inside the inlet tube of the instrument.* It as measured by a Heise Model 623 pressure sensor for the 3760A, and internally by the 3786 WCN. The measurement is used to convert the measured mass flow (FCN or XICN) to volumetric flow and to convert measured particle concentration to equivalent ambient concentration.  

#### CN Counter Inlet Temperature ($^\circ\mathrm{C}$): CNTEMP, TEMP1, TEMP2 {-#cntemp}

*The sample air temperature measured at the intake of the 3760A or within the 3786.* The value is used to convert the measured mass flow (FCN or XICN) to true volumetric flow and to convert measured particle concentration to equivalent ambient concentration.</p>

#### Raw and Corrected CN Counter Sample Flow Rate (SLPM, VLPM): FCN, FCNC {-#fcnc}

*The raw and corrected sample flows in the CN counters* are treated differently for the two models of CN counter. In the 3760A, FCN is measured in standard liters per minute (SLPM) with a mass flow meter. The flow meter gives the volumetric flow rate that would apply under standard conditions of 1013.25 hPa and $21^\circ\mathrm{C}$. FCNC is the corrected sample flow rate in volumetric liters per minute (VLPM) *at instrument pressure and temperature*. **For the 3760A:**    
```{r child='FCNCbox.Rmd'}
```
**In the 3786**, flows are determined in volumetric $\mathrm{cm}^3\mathrm{min}^{-1}$ from the pressure drop maintained across an orifice. The 3786 firmware makes density corrections internally, so its reported sample flow is brought directly into the variable FCNC in units of VLPM.  

#### Raw and Corrected CN Isokinetic Side Flow Rate (SLPM, VLPM): XICN,  XICNC {-#xicnc}

*XICN is the raw isokinetic side flow rate in standard liters per minute (SLPM) measured with a mass flow meter, and XICNC is that flow corrected for pressure and temperature to be the true volumetric flow.* The side flow is adjusted for isokinetic sampling at the inlet, but it is not used further in processing.  
```{r child='XICNCbox.Rmd'}
```

#### CN Counter Output (counts per sample interval): CNTS, CNT1, CNCNTS {-#cnts}

*The raw output count from a condensation nucleus counter.* For the 3760A condensation nucleus counter, the project-dependent sample rate may be chosen in the range from 1–50 Hz but it is typically 10 Hz. In some unusual cases the counts are divided by a selected power of two to keep the counter from overflowing; see project documentation. The 3786 WCN may be programmed to report data at intervals from 0.1–3600 s.<span id="punch7.1" label="punch7.1">`r ifelse(PToDo,"[ToDo7.1]","")`</span></p>

#### Condensation Nucleus (CN) Concentration (cm^-3^): CONCN, CNCONC, CONCNI {-#concn}

*The number concentration of condensation nuclei* in units of particles per cm^3^ *in the ambient air* at flight level. The calculation leading to CONCN includes two corrections. The first accounts for coincidence of particles in the viewing volume at high concentrations and is handled differently in the two types of CN counter. Three instruments have produced this measurement, a TSI 3760A, a TSI 3786, and an experimental water-based CN counter that, in test mode, produced the variable CONCNI. For the 3760A, a statistical adjustment is made based on the average time of a particle in the viewing volume. This correction increases from about 1% at a total concentration of 10^3^cm^-3^ to nearly 11% at 10^4^cm^-3^, but for measurements above about 2x10^4^cm^-3^ significant uncertainty remains because the validity of the coincidence correction is not reliable. The 3786 instead measures the time each detected particle occupies the viewing volume, and this accumulated “dead time” in each sampling interval is subtracted from the elapsed time yielding a “live time” for the determination of sample volume. With this correction an accuracy of 12%, not including statistical counting error, is specified by the manufacturer at concentrations up to 10^5^cm^-3^. The second correction, applied to all CN counters, is a conversion from instrument to ambient conditions.^[Prior to Dec. 2007 the conversion to ambient concentration was not made and concentration was reported for instrument conditions.]
In the following formulae, the corrected flow FCNC in VLPM is explicitly converted to cm^3^s^-1^ by the factor (1000/60).  

**For the 3760A:**  
```{r child='CONCNbox.Rmd'}
```

**For the 3786 WCN:**  
```{r child='CONCN2box.Rmd'}
```

<!-- <h2 id="aerosol-spectrometers">Aerosol Spectrometers</h2> -->
## Aerosol Spectrometers {#aerosol-spec}

For size-resolved measurements of the concentration of aerosol particles, RAF deploys two instruments. The [Ultra High Sensitivity Aerosol Spectrometer](https://www.eol.ucar.edu/instruments/ultra-high-sensitivity-aerosol-spectrometer) (UHSAS) sizes particles in 99 bins from 0.06 to 1.0 $\mu\mathrm{m}$ diameter, and the [Passive Cavity Aerosol Spectrometer Probe](http://www.eol.ucar.edu/instruments/passive-cavity-aerosol-spectrometer-probe) (PCASP) has 30 channels covering the diameter range 0.1 to 3 $\mu\mathrm{m}$. Flow and total concentration variables for these instruments are described in this section, while additional variables are covered along with other 1-D probes in Sect.\ \@ref(sensors-1-D-probes), “Size Distributions (1-D Probes).”  

#### UHSAS Absolute Pressure in Optics Block (kPa): UPRESS {-#upress}

*The pressure internal to the UHSAS instrument.* This is an analog measurement with calibration coefficients as recorded in the attributes for the variable.  

#### Measured and Corrected Sample Flow Rates (units, USMPFLW and PFLW: cm^3^min^-1^; UFLWC and PFLWC: ambient cm^3^s^-1^): USMPFLW, PFLW, UFLWC, PFLWC  {-#pflw}

Unlike the other 1-d probes, both UHSAS and PCASP have internal pumps so their sample volumes are determined from the measured flows and do not depend on true air speed. The UHSAS measures volumetric flow directly, as USMPFLW, and it is adjusted to ambient conditions as UFLWC for the calculation of ambient concentration. The PCASP returns a mass flow referenced to standard conditions (PFLW) and this also is converted to equivalent ambient volumetric flow (PFLWC).  
```{r child='PFLWCbox.Rmd'}
```

#### Total particle counts per sample interval, UHSAS or PCASP: TCNTU, TCNTP {-#tcntu-tcntp}

*The total particle counts in each sample interval* for, respectively, the UHSAS and PCASP instruments. These values are the sum of counts in all cells of the spectrometers, as represented in the vector variables AUHSAS or AS200. See the [discussion of these variables](./cloud-physics-variables.html#tcnt) in Sect.\ \@ref(sensors-1-D-probes). The UHSAS also provided a total count, UACC ("UHSAS Accumulation"). 

#### Concentration, sum over all channels (cm^-3^s^-1^): CONCU, CONCP, CONCU100, CONCU500 {-#concu-concp}

*The particle concentrations summed over all or a subset of channels.* CONCU and CONCP are summed over all channels in the UHSAS and PCASP, respectively, and are calculated as in the following boxed equations. CONCU100 and CONCU500 are concentrations summed over channels in the UHSAS giving particle concentrations for diameters greater than or equal to 100 nm and 500 nm, respectively, and are calculated as for CONCU except with TCNTU replaced by the sum over the appropriate channels.  
```{r child='CONCUbox.Rmd'}
```

<!-- <h2 id="special-aerosol-measurements">Special Aerosol Measurements</h2> -->
## Special Aerosol Measurements {#special-aerosol}

Data from an aerosol mass spectrometer, a scanning mobility particle spectrometer, and a giant nucleus impactor are recorded by these instruments in separate data files and are not recorded by the aircraft data system. The <a name=anc_data>ancillary data sets</a> are not merged into the netCDF archives produced by EOL, so the special data files must be used for these measurements. The data formats are described with the instruments at the references given below:  

* **Aerosol Mass Spectrometer (AMS) data files** contain size-segregated chemical composition of non-refractory, submicron aerosol particles. The instrument is described here: [AMS](https://www.eol.ucar.edu/instruments/time-flight-aerosol-mass-spectrometer).  
* **Scanning Mobility Particle Spectrometer (SMPS) files** contain fine particle differential size distributions. The number of channels and covered size range are variable. Diameter ranges from about 7.5 nm up to about 500 nm (pressure-dependent), and 15 size bins are typical. The instrument is described here: [SMPS](https://www.eol.ucar.edu/instruments/scanning-mobility-particle-spectrometer).  
* **Auto-GNI, GNI Giant Nuclei Impactor (GNI) files** contain dry differential particle size distributions. The instrument is described here: [GNI](https://www.eol.ucar.edu/instruments/giant-nuclei-impactor).  

***  
### Footnotes: {.unnumbered}
